Make any stupid plot you want

### 2021-05-03 draft/sections/math.tex
reworking q and q hat

\subsubsection{Assembly \vmark}
Given the continuities described in \ref{fig:math:graphic:retraction:map}, 
we illustrate a minimal Q that will generate the most minimal visualizations associated with those continuities: non-overlapping scatter points, a non-infinitely thin line, and an image. 
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/scatter_with_s.png}
    \caption{The data is discrete points (precipitation, time). Via \vchannel\ these are converted to (xpos, ypos) and pulled over discrete \gbase. These values are then used to parameterize \gsection\ which returns a color based on the parameters (xpos,ypos) and position $\alpha, \beta$ on $\gbase_{\dbasepoint}$ that \gsection\ is evaluated on. 
    }
    \label{fig:math:artist:q:scatter}
\end{figure}
The scatter plot in \autoref{fig:math:artist:q:scatter} has a constant \textit{size} and color $\gsection_{RGB} = (0,0,0)$ that are defined as part of the point assembly function $\vmark(xpos, ypos)(\alpha, \beta)$.

\begin{figure} [H]
\begin{minipage}{.5\textwidth}
    \centering
    \includegraphics{figures/math/scatter_s.png}
\end{minipage}
\begin{minipage}{.5\textwidth}
    \begin{align*}
        x &= size *\alpha \cos(\beta) + xpos \\[.5\baselineskip]
        y &= size *\alpha \sin(\beta) + ypos
    \end{align*}
\end{minipage}
\label{fig:math:artist:q:scatter:s}
\caption{lookup for scattter glyphs}
\end{figure}

The position of this swatch of color can be computed relative to the location on the disc \((\alpha, \beta) \in \gbase_{\dbasepoint}\) as shown in \autoref{fig:math:artist:q:scatter:s}. \note{EXPLAIN WHAT IS HAPPING IN THE FIGURE THAT GETS BACK TO gsection}

$\gsection(\gbasepoint) = (x, y, 0, 0, 0)$ colors the point (x,y) black.

\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/line_with_s.png}
    \caption{The line fiber $(time,\, temp)$ is thickened with the derivative $(time^{\prime},\, precipitation^{\prime}$ because that information will be necessary to figure out the tangent to the point to draw a line. This is because the line needs to be pushed perpendicular to the tangent of (xpos, ypos). The data is converted to visual characteristics (xpos, ypos). The $\alpha$ coordinates on \gbase\ specifies the position of the line, the $\beta$ coordinate specifies thickness.}
    \label{fig:math:artist:q:line}
\end{figure}
In contrast, the line plot $\vmark(xpos, \hat{n}_{1}, ypos, \hat{n}_{2})(\alpha, \beta)$ in \autoref{fig:math:artist:q:line} has a \vindex\ function that is not only parameterized on \dbasepoint\ but also on the $\alpha$ distance along \dbasepoint\ and corresponding region in \gbase.

\begin{figure} [H]
    \begin{minipage}{.5\textwidth}
        \centering
        \includegraphics{figures/math/line_s.png}
    \end{minipage}
    \begin{minipage}{.5\textwidth}
        \begin{align*}
            \lvert n \rvert &=\sqrt{{n_{1}}^2 + {n_{2}}^2},\,
            \hat{n}_{1} = \frac{n_1}{\lvert n \rvert}, \,\hat{n}_{2} = \frac{n_2}{\lvert n \rvert} \\[.5\baselineskip]
            x &= xpos(\vindex(\alpha)) + width*\beta\hat{n}_1(\vindex(\alpha))  \\[.5\baselineskip]
            y &= ypos(\vindex(\alpha)) + width*\beta\hat{n}_2(\vindex(\alpha)) 
           \end{align*}
    \end{minipage}
    \label{fig:math:artist:q:line:s}
    \caption{lookup for line glyphs}
\end{figure}

As shown in \autoref{fig:math:artist:q:line:s}, line needs to know the tangent of the data to draw an envelope above and below each (xpos,ypos) such that the line appears to have a thickness; therefore the artist takes as input the jet bundle \cite{JetBundle2020,musilovaCalculusVariationsJet2016} $\mathcal{J}^{2}(\dtotal)$ which is the data \dtotal\ and the first and second derivatives of \dtotal. The indexing map $\vindex(\alpha)$ finds the point in \dbase\ corresponding to the region in \gbase\ at coordinate $\alpha$. The section \dsection\ on the \dbasepoint\ that corresponds to the region in \gbase\ returns  the position \textit{xpos, ypos} and the derivatives $\hat{n}_1, \hat{n}_2$ . The derivatives are then multiplied by a \textit{width} parameter to specify the thickness. \note{lil bit messy}

\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/image.png}
    \caption{The only visual parameter an image requires is color since \vindex\ encodes the mapping between position in data and position in graphic. }
    \label{fig:math:artist:q:heatmap}
\end{figure}
In \autoref{fig:math:artist:q:heatmap}, the image 
\begin{equation}
\vmark(xpos, ypos, color)
\label{eq:math:artist:q:image}
\end{equation}
 is a direct lookup into  $\vindex:\gbase\rightarrow\dbase$. The indexing variables $(\alpha, \beta)$ define the distance along the space, which is then used by \vindex\ to map into \dbase\ to lookup the color values 
\begin{equation*}
R = R(\vindex(\alpha, \beta)), \; G = G(\vindex(\alpha, \beta)), \; B = B(\vindex(\alpha, \beta))
\end{equation*}
In the case of an image, the indexing mapper \vindex\ may do some translating to a convention expected by \vmark, for example reorientng the array such that the first row in the data is at the bottom of the graphic.

@@

The graphic base space \gbase\ is not accessible in many architectures, including Matplotlib; instead we can construct a factory function \vmarkd\ over \dbase\ that can build a \vmark. As shown in \autoref{eq:math:artist:diagram}, \vmark\ is a bundle map $\vmark: \vtotalpull\rightarrow \gtotal$ where \vtotalpull\ and \gtotal\ are both bundles over \gbase.

\begin{equation}
    \begin{tikzcd}
        V \arrow[d, "\pi"']             & \xi^{*}V \arrow[d, "\pi"'] \arrow[l, "\xi^{*}"']         \\
        K \arrow[u, "\mu"', bend right] & S \arrow[l, "\xi"'] \arrow[u, "\xi^{*}\mu"', bend right]
    \end{tikzcd}
    \label{eq:math:artist:qhat}
\end{equation}

The preimage of the continuity map $\vpreimg \subset \gbase$ is such that many graphic continuity points $\gbasepoint \in \gbase_{\dbase}$ go to one data continuity point \dbasepoint; therefore, by definition the pull back of \vsection\
\begin{equation}
    \vtotalpull \mid_{\vpreimg} = \vpreimg \times \vfiber
\end{equation}
copies the visual fiber \vfiber\ over the the points \gbasepoint\ in graphic space \gbase\ that correspond to one \dbasepoint\ in data space \dbase. This set of points \gbasepoint\ are the preimage \vpreimg\ of \dbasepoint.

>>>>

\subsubsection{Assembly \vmark}
Given the continuities described in \ref{fig:math:graphic:retraction:map}, 
we illustrate a minimal Q that will generate the most minimal visualizations associated with those continuities: non-overlapping scatter points, a non-infinitely thin line, and an image. 
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/scatter_with_s.png}
    \caption{The data is discrete points (precipitation, time). Via \vchannel\ these are converted to (xpos, ypos) and pulled over discrete \gbase. These values are then used to parameterize \gsection\ which returns a color based on the parameters (xpos,ypos) and position $\alpha, \beta$ on $\gbase_{\dbasepoint}$ that \gsection\ is evaluated on. 
    }
    \label{fig:math:artist:q:scatter}
\end{figure}
The scatter plot in \autoref{fig:math:artist:q:scatter} has a constant size and color $\gsection_{RGB} = (0,0,0)$ that are defined as part of the point assembly function.

\begin{figure} [H]

\begin{minipage}{.5\textwidth}
    \begingroup \leqnomode
    \begin{equation}
        \vmark(xpos, ypos)(\alpha, \beta)
        \label{eq:eq:math:artist:q:scatter}
    \end{equation}
    \endgroup
    \begin{align*}
        x &= \text{size} *\alpha \cos(\beta) + xpos \\[.5\baselineskip]
        y &= \text{size} *\alpha \sin(\beta) + ypos
    \end{align*}
\end{minipage}
\begin{minipage}{.5\textwidth}
    \centering
    \includegraphics{figures/math/scatter_s.png}
\end{minipage}
\caption{The simplest form of the scatter plot takes as input the expected position of the marker in visual space \textit{(xpos, ypos)}. The marker shape is determined by the polar coordinates $(\alpha, \beta)$ on the disc; these coordinates dictate whether anything is drawn at that region of \gbase. To obtain the color of the pixel at \textit{(x,y)}, the region on \gbase\ is scaled by a constant size and shifted by the \textit{xpos} and \textit{ypos}.}
\label{fig:math:artist:q:scatter:s}
\end{figure}
The position of this swatch of color is computed relative to the location on the disc \((\alpha, \beta) \in \gbase_{\dbasepoint}\) as shown in \autoref{fig:math:artist:q:scatter:s}. The region $\alpha, \beta)$ is scaled by a constant size and shifted by \textit{xpos} and \textit{ypos}. This computation yields the values \textit{(x,y)} that map into \gfiber\ and have a corresponding function $\gsection(\gbasepoint) = (x, y, 0, 0, 0)$ which colors the point \textit{(x,y)} black.

\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/line_with_s.png}
    \caption{The line fiber $(time,\, temp)$ is thickened with the derivative $(time^{\prime},\, precipitation^{\prime}$ because that information will be necessary to figure out the tangent to the point to draw a line. This is because the line needs to be pushed perpendicular to the tangent of (xpos, ypos). The data is converted to visual characteristics (xpos, ypos). The $\alpha$ coordinates on \gbase\ specifies the position of the line, the $\beta$ coordinate specifies thickness.}
    \label{fig:math:artist:q:line}
\end{figure}
In contrast, the line plot in \autoref{fig:math:artist:q:line} has a \vindex\ function that is not only parameterized on \dbasepoint\ but also on the $\alpha$ distance along \dbasepoint\ and corresponding region in \gbase.

\begin{figure} [H]
    \begin{minipage}{.5\textwidth}
        \begingroup \leqnomode
        \begin{equation}
            \vmark(xpos, n_1, ypos, n_2 )(\alpha, \beta) 
            \label{eq:eq:math:artist:q:line}
        \end{equation} \endgroup
        \begin{align*}
            \lvert n \rvert &=\sqrt{{n_{1}}^2(\vindex(\alpha)) + {n_{2}}^2(\vindex(\alpha))} \\
            \hat{n}_{1} &= \frac{n_1(\vindex(\alpha))}{\lvert n \rvert}, \,\hat{n}_{2} = \frac{n_2(\vindex(\alpha))}{\lvert n \rvert} \\[.5\baselineskip]
            x &= xpos(\vindex(\alpha)) + \text{width}*\beta\hat{n}_1(\vindex(\alpha))  \\
            y &= ypos(\vindex(\alpha)) + \text{width}*\beta\hat{n}_2(\vindex(\alpha)) 
           \end{align*}
    \end{minipage}
    \begin{minipage}{.5\textwidth}
        \centering
        \includegraphics{figures/math/line_s.png}
    \end{minipage}
    \caption{The \textit{xpos} and \textit{ypos} variables give the position of the line in screen space, but render an infinitely thin line. To draw equidistant lines parallel to(\textit{xpos}, \textit{ypos}), defined by the distance  $(n_1, n_2)$, requires the derivatives $(n_{1} = xpos^{\prime}, n_{2} = ypos^{\prime})$. The position (\textit{xpos}, \textit{ypos}) and width of the line is then used to determine whether a pixel is colored at the position (\textit{x}, \textit{y}). The values in data space are only looked up via the $\alpha$ coordinate of \gbase\ because it maps to a location on \dbase. The $\beta$ parameter is used to specify how thick the line is in conjunction with the constant width. }
    \label{fig:math:artist:q:line:s}
\end{figure}
As shown in \autoref{fig:math:artist:q:line:s}, line needs to know the tangent of the data to draw an envelope above and below each (xpos,ypos) such that the line appears to have a thickness; therefore the artist takes as input the jet bundle \cite{JetBundle2020,musilovaCalculusVariationsJet2016} $\mathcal{J}^{2}(\dtotal)$ which is the data \dtotal\ and the first and second derivatives of \dtotal. The indexing map $\vindex(\alpha)$ finds the point in \dbase\ corresponding to the region in \gbase\ at coordinate $\alpha$. The section \dsection\ on the \dbasepoint\ that corresponds to the region in \gbase\ returns the position \textit{xpos, ypos} and the derivatives $\hat{n}_1, \hat{n}_2$ . The derivatives are then multiplied by a width parameter to specify the thickness of the line. This is then used to determine the color of the pixel at \textit{(x,y)}.

\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/image.png}
    \caption{Via \vindex\, the artist maps from a point \textit{(x,y)} on the screen to a corresponding point on \dbase. This maps into \dfiber\ via \dsection. These data points are converted to visual points via \vchannel\, and then \vmark\ assembles the \textit{(xpos, ypos, color)} parameters into attributes of each pixel. }
    \label{fig:math:artist:q:heatmap}
\end{figure}
In \autoref{fig:math:artist:q:heatmap}, the image is a direct lookup into  $\vindex:\gbase\rightarrow\dbase$. The indexing variables $(\alpha, \beta)$ define the distance along the space, which is then used by \vindex\ to map into \dbase\ to lookup the color values.
\begin{equation}
    \vmark(xpos, ypos, color)(\alpha, \beta)
    \label{eq:math:artist:q:image}
    \end{equation}
\begin{align*}
x &= xpos(\vindex(\alpha))\\
y &= ypos(\vindex(\beta))\\
R, G, B &= color(\vindex(\alpha, \beta))
\end{align*}
In the case of an image, the indexing mapper \vindex\ may do some translating to a convention expected by \vmark, for example reorientng the array such that the first row in the data is at the bottom of the graphic.

@@

The graphic base space \gbase\ is not accessible in many architectures, including Matplotlib; instead we can construct a factory function \vmarkd\ over \dbase\ that can build a \vmark. As shown in \autoref{eq:math:artist:diagram}, \vmark\ is a bundle map $\vmark: \vtotalpull\rightarrow \gtotal$ where \vtotalpull\ and \gtotal\ are both bundles over \gbase.
%[column sep=huge]
\begin{equation}
    \begin{tikzcd}
        \textcolor{gray!50}{\dtotal} \arrow[r, "\vchannel", color=gray!50] \arrow[rd, "\pi"', color=gray!50] & \vtotal \arrow[d, "\pi"']                  & \vindex^*\vtotal \arrow[r, "\vmark", color=gray!50] \arrow[d, "\vindex^*\pi"'] \arrow[l,  "\vindex^*"'] & \textcolor{gray!50}{\gtotal} \arrow[ld, "\pi", color=gray!50] \\
                                                          & \dbase \arrow[u, "\vsection"', bend right] & \gbase \arrow[l, "\vindex"'] \arrow[u, "\vindex^*\vsection"', bend right]               &                          
        \end{tikzcd}
        \label{eq:math:artist:qhat}
\end{equation}
The preimage of the continuity map $\vpreimg \subset \gbase$ is such that many graphic continuity points $\gbasepoint \in \gbase_{\dbase}$ go to one data continuity point \dbasepoint; therefore, by definition the pull back of \vsection\
\begin{equation}
    \vtotalpull \mid_{\vpreimg} = \vpreimg \times \vfiber
\end{equation}
copies the visual fiber \vfiber\ over the the points \gbasepoint\ in graphic space \gbase\ that correspond to one \dbasepoint\ in data space \dbase. This set of points \gbasepoint\ are the preimage \vpreimg\ of \dbasepoint.
